Speculation in foreign exchange: noise or information

### 2015-06-07 Knitr/NewPaper.Rnw
NewPaper.Rnw has been taken from the previous commit. It did not seem to have been updated. NOw the file has the methods (though these need to be looked at) and it hs the changes to reason for fx etc.

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculative positions with the movement of prices in the foreign exchange market.   If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.  
    The foreign exchange market is utilised because it is the largest and most liquid market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn \citet{BISFX2013}. In addition, the cost of trading in this markket, when measurd in terms of the size of the bid-ask spread, is extremely low. An analysis of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately)\citep{Steely2013}. The idea of noise trading may be particularly important for the foreign exchange market where it is widely perceived that there are a relatively large number of traders engaged in technical analysis.  See \citep{FrootChartists} for a survey.

@@

\subsection*{The problem of uninformed noise-traders}
The traditional Keynesian view of speculation asserted that it would be an additional noise that would make it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value. This view finds some supported in Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}. In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved. 
%It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
\subsection{Interplay between Informed and uninformed}
\citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''. In this case, the construction olves the problem of financial market trading. If economic agents know the fundamental value of a security how can some be buying and some be selling? There must be different opinions.  Why Will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information? However, with noise-traders, the informed can take advantage of their information.  This creaes a justification for acquiring information. Noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information. The divergence between this price and fundametnal value is essential to encourage te process of price discovery.

Fisher Black identified noise as ``the arbitrary element in expectations'' \citet[p.529]{BlackNoise}. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.

@@

One approach has been to assume that noise-traders utilise simple investment strategies that involve momentum or reversals.  For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.  In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation. Alternatively, speculation can be identified with momentum trading.   The literature on heterogeneous traders, another solution to the justificcation for trading securities, has very frequently used noise-traders as trend chasers or followers of technical analysis.  These models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.

@@

For example, the debate about a tax on financial transactions, \emph{the Tobin tax} \citet{TobinTax}, has been intended to reduce speculation.  The surge in commodities prices that was seen prior to the GFC was seen to be partially the consequence of increased speculation in commodities and the creation of exchange traded funds (ETF) focusing on commodities markets.  This argument depends on the idea that speculators will drive the price away from its fundamental value.   
In addition, so-called \emph{Volcker Rule} would prevent US investment banks trading on their own-account.  This, it has been argued, will reduce liquidity in the bond market by causing investment banks to hold lower levels of bond inventory on their books.  This may increase the amount of \emph{price slippage} when final investor adjust their positions. In addition,

\section{Methodology}

The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculators' activity in the  foreign exchange market .   This will involve three steps:  first,assess the price implication of the noise-trader modele when speculative sentiment or activity becomes extreme; second, measure speculative activity in the foreign exchange market; finally,  whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.

@@

where $\rho^*$ is the average mistake of the noise-traders and $\sigma_{\rho}^2$ is the variance of the misperception.      Assuming an exponential utility function or constant absolute risk aversion\footnote{Though \citep{TobinLiquidity} used a quadratic utility function to show that investor preference could be defined by just the mean and the variance of the returns, the increasing absolute risk aversion this implies does not seem to be a suitable property as it implies people become more unwilling to take risk as they become more wealthy (\citep{Hicks1962} and \citep{ArrowRisk}.   Constant absolute risk aversion is the \citep{PrattCARA}, \citep{ArrowRisk}  measure based on $u(c) = 1 - e^{-\gamma c}$ where the \emph{coefficient of absolute risk aversion} $A(c) = -\frac{u''(c)}{u'(c)} = \gamma$.  This is a standard device.}, each type of investor chooses the proportion $(\lambda)$ of the risky asset to hold to maximise utility.

\begin{equation}
E(U) = \bar{w} - \gamma \sigma_{w}^2
\end{equation}

where $\bar{w}$ is end of period wealth, $\gamma$ is a measure of absolute risk aversion and $\sigma_w^2$ is the variance of wealth. Utility is a function of risk and return.  This depends on holdings of the risky asset.   For the informed trader this is,

\begin{equation}
E(U) = c_0 + \lambda_t^i[r + _tp_{t+1} - p_t(1 + r)] - \gamma (\lambda_t^i)^2 \left (_t\sigma^2_{p, t+1} \right ) 
\end{equation}

where $c_0$ is the endowment, following the notation in the original paper, $_tp_{t+1}$ is the expected price of the risky asset at $t + 1$, conditional on information at time $t$ and $_t\sigma^2_{\rho, t+1}$ is the expected variance of the misperception conditional on information at time $t$.

@@

The paper makes use of the utility function to compare the outcomes for the two types of investor for different realisations of the parameters.  They note four effects:  the \emph{Friedman Effect}, which is the tendency of informed investors to push prices back towards fundamental value; \emph{Price Pressure} as the way that speculative noise-traders buy more and drive the price higher as they become more bullish; \emph{Hold More},whereby the noise-traders hold more of the risky asset; \emph{Create Space} is the added risk that is caused by an increase in the variability of noise-trader belief.  The Friedman Effect and Price Pressure tend to reduce noise-trader utility as they work to create losses or reduced gains for noise-trader positions; Hold More and Create Space can increase noise-trader utility and profits \citep[pp. 14-15]{Delong1990noise}. 
%Put something on these four in the conclusion. 
A number of important ideas flow from the model.  The main findings are that noise-traders will not necessarily disappear as Friedman had suggested \citep{FriedmanPositive}.  While noise-traders will tend to get less utility than sophisticated traders, this does not necessarily drive them from the market.  The noise-traders expected to get higher utility due to their misperception.  The reduction in utility comes mainly from the fact that they are not compensated for taking risk that they are not aware of.  This is consistent with the type of over-confident, entrepreneurial misperception discussed in Section \ref{sec:prob} in relation to the returns to speculation.  If the noise-traders become aware of the risk that they are taking they are less likely to undertake the activity.

@@

%One of the key assumptions of the model is the two period construction that means that final value is not determined.  This means that informed investors cannot just wait for the fundamental value to assert itself.  Therefore, the model is consistent with the sort of myopia that may be caused by impatience \citep{Keynes1936}, funding costs, social pressure and institutional features. % references needed.  
%\footnote{An good practical example is that of Tony Dye the Philips and Drew Fund Manager who refused to participate in the technology boom of the late 1990s.  The relative under-performance of the fund cause a loss of investors and his eventual resignation in March 2000.  This is the same date as the peak of the US technology index NASDAQ.  It is also consistent with the words of Citicorp Chief Executive Charles Prince who is reported to have said, "As long as the music's playing, you've got to get up and dance" when asked by the \emph{Financial Times} in July 2007 about the bank's lending activity.}

%is this paragraph necessary?
%The De Long and others do not go into the details of the misperception that is suffered by the noise traders.  There are a number of explanation for this deviation of the misperception from zero.  The first are the behavioural biases that were addressed in section \ref{sec:behavioural}.   If investors generally are suffering from the \emph{representativeness} heuristic an over-focus on the immediate information that the expense of the bigger picture or the base risk to securities value would cause a systematic deviation in favour or optimism or pessimism.  This effect can be exacerbated by the overconfidence that has been identified generally in the behavioural economic literature and by Odean specifically for investment markets.  Systematic bias creates price pressure which feeds back into a successful outcome that is regarded as being the consequence of skill, ability and good judgment and investment talent rather than luck and temporary circumstance.  %more may be required here Hirshliefer and Luo (2001) on the survival of noise traders.
%There is also the social pressure and a feedback mechanism that comes from taking an action that others are taking.  Therefore, if others have the same investment strategy, this can make investors more confident and, in some cases, will cause investors to infer private information from the actions of other investors who are basing investment decisions on the same style of imitation.  All the investors think that they are in good company and that the knowledge of the others fills the gaps where their own uncertainty begins. % need references for this.

\
\bibliography{../../../../myrefs}

>>>>

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculative positions with the movement of prices in the foreign exchange market.   If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.  
    The foreign exchange market is utilised because it is the largest and most liquid market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn \citet{BISFX2013}. In addition, the cost of trading in this markket, when measurd in terms of the size of the bid-ask spread, is extremely low. An analysis of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately)\citep{Steely2013}.

@@

\subsection*{The problem of uninformed noise-traders}
The traditional Keynesian view of speculation asserted that it would be an additional noise that would make it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value. This view finds some supported in Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}. In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved. 
%It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
\subsection{Interplay between Informed and uninformed}
\citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''. In this case, the construction olves the problem of financial market trading. If economic agents know the fundamental value of a security how can some be buying and some be selling? There must be different opinions.  Why Will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information? However, with noise-traders, the informed can take advantage of their information.  This creaes a justification for acquiring information. Noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information. The divergence between this price and fundametnal value is essential to encourage te process of price discovery. 
% What where the reasons for creating the concept of noise-traders?  There are three:  the rtionale for trading, the motivtion or mraket-marking and... 
Fisher Black identified noise as ``the arbitrary element in expectations'' \citet[p.529]{BlackNoise}. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.

@@

One approach has been to assume that noise-traders utilise simple investment strategies that involve momentum or reversals.  For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.  In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation. Alternatively, speculation can be identified with momentum trading.   The literature on heterogeneous traders, another way that the trading securities is explained, has very frequently used noise-traders as trend-chasers or followers of technical analysis.  These models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.

@@

For example, the debate about a tax on financial transactions, \emph{the Tobin tax} \citet{TobinTax}, has been intended to reduce speculation.  The surge in commodities prices that was seen prior to the GFC was seen to be partially the consequence of increased speculation in commodities and the creation of exchange traded funds (ETF) focusing on commodities markets.  This argument depends on the idea that speculators will drive the price away from its fundamental value.

In addition, so-called \emph{Volcker Rule} would prevent US investment banks trading on their own-account.  This, it has been argued, will reduce liquidity in the bond market by causing investment banks to hold lower levels of bond inventory on their books.  This may increase the amount of \emph{price slippage} when final investor adjust their positions. In addition,

\section{Methodology}
The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculator activity in the  foreign exchange market. This will be use to gain a fuller understanding of speculative behaviour.   This will involve three steps:  first,assess the price implication of the noise-trader model;  second, measure speculative activity in the foreign exchange market; finally, assess whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.  If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.  
% where it is widely perceived that there are a relatively large number of traders engaged in technical analysis.  See \citep{FrootChartists} for a survey.  %this may go in the explanation for why fx is used.  Use this elsewhere. Add more?

@@

@@

The paper makes use of the utility function to compare the outcomes for the two types of investor for different realisations of the parameters.  They note four effects:  the \emph{Friedman Effect}, which is the tendency of informed investors to push prices back towards fundamental value; \emph{Price Pressure} as the way that speculative noise-traders buy more and drive the price higher as they become more bullish; \emph{Hold More},whereby the noise-traders hold more of the risky asset; \emph{Create Space} is the added risk that is caused by an increase in the variability of noise-trader belief.  The Friedman Effect and Price Pressure tend to reduce noise-trader utility as they work to create losses or reduced gains for noise-trader positions; Hold More and Create Space can increase noise-trader utility and profits \citep[pp. 14-15]{Delong1990noise}. 
%Put something on these four in the conclusion. 
A number of important ideas flow from the model.  The main findings are that noise-traders will not necessarily disappear as Friedman had suggested \citep{FriedmanPositive}.  While noise-traders will tend to get less utility than sophisticated traders, this does not necessarily drive them from the market.  The noise-traders expected to get higher utility due to their misperception.  The reduction in utility comes mainly from the fact that they are not compensated for taking risk that they are not aware of.  This is consistent with the type of over-confidence that has been found in day-traders \citet{odean1998volume}

@@

\bibliography{../../../../myrefs}
